There once was a grid at ol' Carkeek

### 2016-08-13 Manuscript/v0.0/carkeek_grid.tex
decontamination methods

@@

%METHODS
%
%
%Reads that contain primer index sequences (or combinations thereof) that we did not actually use in laboratory procedures are known to ?(be derived from contamination) be a result of processes that leave the same signature as contamination. For example, errors during oligonucleotide synthesis of the indexed primers would cause reads to be erroneously assigned to samples.
%Because this represents the minimum rate at which we can be certain that sequences from one environmental sample could be erroneously assigned, we considered for further analysis only those reads occurring with greater frequency than this.
%
We calculated rates of cross-library contamination by counting occurrences of primer sequences: 12S primer sequences appearing in a 16S library (and vice versa) indicate an error in the preparation or sequencing procedures.

@@

%To infer a general estimate of the rate of cross-sample contamination likely to be associated with lab procedures, we calculated the proportion of reads containing primer index sequences (or combinations thereof?) that we did not actually use
%
%RESULTS
%Reads were of generally high-quality, with ?(96.48)? % of base calls having a 0.001 or lower probability of incorrect base call (Phred q-score of 30 or higher).
%
%
%Total sequencing clusters passing filter per read: 371,576,190 on NEXTSEQ
%
% THESE BELOW ARE FROM A MISEQ RUN
%reads, reads passing filter, % >= Q30, for each run ( 7,052,307 ; 6,752,829; 97.47 :: 7,052,307 ; 6,752,829 ; 95.50)
%
%Cluster density of 358 \UTF{00EF}\UTF{00BF}\UTF{0153}22	clusters per mm^2
%
%XXX % of reads aligned to the PhiX genome
%(31.02 \UTF{00EF}\UTF{00BF}\UTF{0153}0.59	; 29.17 \UTF{00EF}\UTF{00BF}\UTF{0153}0.89)
%
%and error rate as determined from alignment of PhiX reads to the PhiX genome was low (Read 1: 1.26 \UTF{00EF}\UTF{00BF}\UTF{0153}0.07; Read 2: 1.38 \UTF{00EF}\UTF{00BF}\UTF{0153}0.07)
%----------------------------------------------------------------------------

>>>>

Reads that do not contain primer index sequences (or combinations thereof) in the expected position are known to ?(be derived from contamination | result from processes) that leave the same signature as contamination.
For example, errors during oligonucleotide synthesis or sequencing of the indexed primers could cause reads to be erroneously assigned to samples.
Similarly, we calculated rates of cross-library contamination by counting occurrences of primer sequences from 12S amplicons prepared separately but pooled and sequenced alongside our samples.
These occurrences indicate an error in the preparation or sequencing procedures.
This represents a general minimum rate at which we can expect that sequences from one environmental sample could be erroneously assigned, and so we considered for further analysis only those reads occurring with greater frequency than this across the entire dataset.

@@

@@

%RESULTS
%Reads were of generally high-quality, with XXX% of base calls having a 0.001 or lower probability of incorrect base call (Phred q-score of 30 or higher).
%Total sequencing clusters passing filter per read: 371,576,190 on NEXTSEQ
%
% THESE BELOW ARE FROM A MISEQ RUN
% reads passing filter, % >= Q30, for each run ( 7,052,307 ; 6,752,829; 97.47 :: 7,052,307 ; 6,752,829 ; 95.50)
%
%Cluster density of 358 \UTF{00EF}\UTF{00BF}\UTF{0153}22	clusters per mm^2
%
%XXX % of reads aligned to the PhiX genome
%(31.02 \UTF{00EF}\UTF{00BF}\UTF{0153}0.59	; 29.17 \UTF{00EF}\UTF{00BF}\UTF{0153}0.89)
%
%and error rate as determined from alignment of PhiX reads to the PhiX genome was low (Read 1: 1.26 \UTF{00EF}\UTF{00BF}\UTF{0153}0.07; Read 2: 1.38 \UTF{00EF}\UTF{00BF}\UTF{0153}0.07)
%----------------------------------------------------------------------------
